---
title: Final Activity
format: html
author: Dominic Sitto, Xinyi Lu, Jinman Zhang
date: December 4th, 2024
self-contained: true
---

## dependencies

```{r}
library(ggplot2)

library(tidyverse)
library(Stat2Data)
library(dplyr)
library(ggplot2)
library(dcData)

foodDF <- read_csv("food-world-cup-data.csv")
```

### 1. The file contains abnormal characters such as Ê, clean up these characters.

--\>I replaced it in Excel.

## Adding titles to columns
```{r}
colnames(foodDF) <- c("RespondentID","Generally speaking, how would you rate your level of knowledge of cuisines from different parts of the world?","How much, if at all, are you interested in cuisines from different parts of the world?","Please rate how much you like the traditional cuisine of Algeria.","Please rate how much you like the traditional cuisine of Argentina.","Please rate how much you like the traditional cuisine of Australia.","Please rate how much you like the traditional cuisine of Belgium.","Please rate how much you like the traditional cuisine of Bosnia and Herzegovina.","Please rate how much you like the traditional cuisine of Brazil.","Please rate how much you like the traditional cuisine of Cameroon.","Please rate how much you like the traditional cuisine of Chile.","Please rate how much you like the traditional cuisine of Colombia.","Please rate how much you like the traditional cuisine of Costa Rica.","Please rate how much you like the traditional cuisine of Croatia.","Please rate how much you like the traditional cuisine of Ecuador.","Please rate how much you like the traditional cuisine of England.","Please rate how much you like the traditional cuisine of France.","Please rate how much you like the traditional cuisine of Germany.","Please rate how much you like the traditional cuisine of Ghana.","Please rate how much you like the traditional cuisine of Greece.","Please rate how much you like the traditional cuisine of Honduras.","Please rate how much you like the traditional cuisine of Iran.","Please rate how much you like the traditional cuisine of Italy.","Please rate how much you like the traditional cuisine of Ivory Coast.","Please rate how much you like the traditional cuisine of Japan.","Please rate how much you like the traditional cuisine of Mexico.","Please rate how much you like the traditional cuisine of the Netherlands.","Please rate how much you like the traditional cuisine of Nigeria.","Please rate how much you like the traditional cuisine of Portugal.","Please rate how much you like the traditional cuisine of Russia.","Please rate how much you like the traditional cuisine of South Korea.","Please rate how much you like the traditional cuisine of Spain.","Please rate how much you like the traditional cuisine of Switzerland.","Please rate how much you like the traditional cuisine of United States.","Please rate how much you like the traditional cuisine of Uruguay.","Please rate how much you like the traditional cuisine of China.","Please rate how much you like the traditional cuisine of India.","Please rate how much you like the traditional cuisine of Thailand.","Please rate how much you like the traditional cuisine of Turkey.","Please rate how much you like the traditional cuisine of Cuba.","Please rate how much you like the traditional cuisine of Ethiopia.","Please rate how much you like the traditional cuisine of Vietnam.","Please rate how much you like the traditional cuisine of Ireland.","Gender","Age","Household Income","Education","Location (Census Region)"
)
```

### 2. Data impute

```{r}

# Impute data: first impute by group, then use global median to fill remaining missing values
foodDF_imputed <- foodDF %>%
  # Group by Location, Gender, and Age, and impute missing values with the group median
  group_by(`Location (Census Region)`, Gender, Age) %>%
  mutate(across(where(is.numeric), ~ ifelse(is.na(.), median(., na.rm = TRUE), .))) %>%
  ungroup() %>%
  # For any remaining missing values, impute them with the global median
  mutate(across(where(is.numeric), ~ ifelse(is.na(.), median(., na.rm = TRUE), .)))

# View the imputed data
print(head(foodDF_imputed))

# Save the imputed data to a file
write_csv(foodDF_imputed, "cleaned_food_world_cup_data_filled_by_global_median.csv")

```

### **Advantages of this Method**

-   **Balances both local and global characteristics**: First, impute by group to capture the characteristics within each group; then use the global median to fill in missing values that could not be imputed by the group, ensuring that all data is completed.

-   **Wide applicability**: Even if all values within a group are missing, the imputation will not fail because the global median serves as a fallback solution.
