---
title: Final Activity
format: html
author: Dominic Sitto, Xinyi Lu, Jinman Zhang
date: December 4th, 2024
self-contained: true
---

## dependencies

```{r}
library(ggplot2)

library(tidyverse)
library(Stat2Data)
library(dplyr)
library(ggplot2)
library(dcData)

tacoData <- read_csv("food-world-cup-data.csv")
```

### 1. The file contains abnormal characters such as Ê, clean up these characters.

--\>I replaced it in Excel.

### 2. Data impute

```{r}

# Impute data: first impute by group, then use global median to fill remaining missing values
tacoData_imputed <- tacoData %>%
  # Group by Location, Gender, and Age, and impute missing values with the group median
  group_by(`Location (Census Region)`, Gender, Age) %>%
  mutate(across(where(is.numeric), ~ ifelse(is.na(.), median(., na.rm = TRUE), .))) %>%
  ungroup() %>%
  # For any remaining missing values, impute them with the global median
  mutate(across(where(is.numeric), ~ ifelse(is.na(.), median(., na.rm = TRUE), .)))

# View the imputed data
print(head(tacoData_imputed))

# Save the imputed data to a file
write_csv(tacoData_imputed, "cleaned_food_world_cup_data_filled_by_global_median.csv")

```

### **Advantages of this Method**

-   **Balances both local and global characteristics**: First, impute by group to capture the characteristics within each group; then use the global median to fill in missing values that could not be imputed by the group, ensuring that all data is completed.

-   **Wide applicability**: Even if all values within a group are missing, the imputation will not fail because the global median serves as a fallback solution.
